### 2022-11-09 Manuscripts/main.tex
Update markdown

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{lineno}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{calrsfs}
\usepackage{pgfplots}
\usepackage[affil-it]{authblk}
\usepackage{color}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{cancel}
\usepackage{manyfoot}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{tcolorbox}
\usepackage[english]{babel}
\usepackage{amsthm}
\usepackage{soul}
\usepackage[charter,cal=cmcal]{mathdesign} % Specify font for serif family
\usepackage{avant} % Specify avant font for sans serif family
\usepackage{mathtools} % Replaces amsmath
\usepackage{sectsty} % Specify section formatting
	\sectionfont{\sffamily\color{DarkBlue}}
	\subsectionfont{\sffamily\color{LightBlue}}
	\subsubsectionfont{\small\sffamily}

\DeclareMathAlphabet{\pazocal}{OMS}{zplm}{m}{n}
\newcommand{\La}{\mathcal{L}}
\newcommand{\Lb}{\pazocal{L}}
\newcommand{\la}{\mathcal{l}}
\newcommand{\lb}{\pazocal{l}}
\newcommand{\M}{\pazocal{M}}
\newcommand{\Ha}{\mathcal{H}}
\newcommand{\Hb}{\pazocal{H}}
\newcommand{\Pa}{\mathcal{P}}
\newcommand{\Pb}{\pazocal{P}}

% Reference
\definecolor{wine_red}{RGB}{178, 8, 56}
% Emphasizing term
\definecolor{lan_red}{RGB}{238, 46, 36}
\definecolor{ReD}{RGB}{255,0,0}
% Emphasizing number
\definecolor{violet}{RGB}{133, 12, 112}
% Comments and footnotes
\definecolor{sea_blue}{RGB}{0, 84, 158}
% Axlers
\definecolor{axler_yellow}{RGB}{255, 250, 198}
\definecolor{background}{RGB}{251, 252, 240}
\definecolor{light_blue}{RGB}{235, 240, 252}

\hypersetup{
    colorlinks = true,
    linkcolor = wine_red
}

\newcommand{\R}{\mathbb R}
\newcommand{\C}{\mathbb C}
\newcommand{\Q}{\mathbb Q}
\newcommand{\F}{\mathbb F}
\newcommand{\N}{\mathbb N}
\newcommand{\Z}{\mathbb Z}
\newcommand{\imply}[0]{\Rightarrow}
\newcommand{\emphasize}[1]{\textcolor{ReD}{\textbf{\hl{#1}}}}
\newcommand{\important}[1]{{\textbf{\hl{#1}}}}
\newcommand{\num}[1]{\textcolor{violet}{\emph{\textbf{#1}}}}
\newcommand{\foot}[1]{{\footnote{\textcolor{sea_blue}{#1}}}}
\newcommand{\leftb}{\left(}
\newcommand{\rightb}{\right)}
\newcommand{\inner}[2]{$\langle${$#1$},{$#2$}$\rangle$}
\newcommand{\innereq}[2]{\langle{#1},{#2}\rangle}
\newcommand{\innerproduct}[1]{$\langle${$#1$}$\rangle$}
\newcommand{\innerproducteq}[1]{\langle{#1}\rangle}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{|#1|}
\newcommand{\braket}[1]{\left\{ {#1} \right\}}
\newcommand{\qedwhite}{\hfill \ensuremath{\Box}}
\newcommand{\compart}{{\color{lan_red}\rule{\linewidth}{0.5mm}}} 
\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\vect}[1]{\left( #1 \right)}

\newcommand{\dimension}{\text{dim}}
\newcommand{\nullspace}{\text{null}}
\newcommand{\range}{\text{range}}
\newcommand{\operator}[1]{#1 \in \Lb(V)}
\newcommand{\operatorspace}[2]{#1 \in \Lb(#2)}
\newcommand{\V}{\R^n}

\pagestyle{fancy}
\fancyhead[L]{Blair Yang}
\fancyhead[R]{Supplementary Materials}

\usepackage{xcolor} % required to defined colours
	\definecolor{DarkBlue}{HTML}{25355A}
	\definecolor{LightBlue}{HTML}{007FA3}
	\definecolor{Yellow}{HTML}{A37500}
	\definecolor{Red}{HTML}{A3002E}
	\definecolor{Skyblue}{HTML}{007FA3}
	\definecolor{Background_red}{HTML}{F6E5EA}
	\definecolor{bgyellow}{HTML}{F2ECD9}
	\definecolor{darkyellow}{HTML}{A07300}
	\definecolor{lan_purple}{HTML}{850C70}

% \usepackage[sfdefault]{carlito}
% \usepackage{fontspec}
% \setmainfont{Arial}

\newcommand{\urlref}{\hl{[REF]}}
\newcommand{\pendingref}{\textbf{\hl{[REF]}}$ \ $}
\newcommand{\unknownref}{\textcolor{ReD}{\textbf{\hl{[REF]}}}$ \ $}
\newcommand{\pendingdata}{\textbf{\hl{$<$DATA$>$}}$ \ $}
\newcommand{\unknowndata}{\textcolor{ReD}{\textbf{\hl{$<$DATA$>$}}}$ \ $}

\title{Supplementary Materials}
\author{Blair Yang}
\date{August 2022}

\begin{document}

\section*{Style}
\urlref $ \ \ \ $Pending reference, with the following URL\\
\pendingref $ \ \ $ Pending reference, with the following URL\\
\unknownref $ \ \ $ Pending reference, with un-specified topic (need to find relevant papers)\\
\pendingdata $ \ \ $ Pending data, need to be filled with data from the cited papers\\
\unknowndata $ \ \ $ Pending data, need to be find relevant paper with such data

\section{Methodology (supplementary material)}
\subsection{Data acquisition}

\subsubsection{Epidemiological data}

We acquired the time-series data of COVID-19 cases, deaths, and vaccination data in the 26 Public health units (PHU) since January 2020 from Ontario public health. \cite{ref1} The original data was stratified into \pendingdata age bands (\pendingdata).

\subsubsection{Vaccine data}

We acquired the provincial time-series vaccination data in Ontario from Ontario public health, covering the percentage population who has taken first does, second dose, and third does. We assumed a uniform distribution of the vaccination population ratio for every PHU.

\subsubsection{Labour force data}

We used the labour force age distribution of Canada from \pendingref \emphasize{(2016 Census)} in year \pendingdata and the employment rate in August 2022 \pendingdata by statistics Canada \pendingref.

\subsubsection{Commute data}

The commuting data in Ontario was excerpted from the 2016 demographic census. We assumed a constant age distribution.

\subsection{Data Calibration}

\subsubsection{Age-specific data calibration}

Due to the different dimensionality of the contact matrix in Canada (16 age-bands with 5-years sensitivity) and the Ontario epidemiological data (6 age-bands with 10-years sensitivity), we augmentation all of the age-specific data using difference of Gaussians (DoG), the estimated percentage population with certain characteristic (i.e, cases, deaths, vaccinated) with an integer-valued age $i$ is estimated to be:
\begin{gather}
    \mu_{i} = \Gamma_{\sigma_1, \sigma} (i) = \frac{e^{-\frac{i^2}{2 \sigma_1^2}}}{\sigma_1 \sqrt{2\pi}} I - \frac{e^{-\frac{i^2}{2 \sigma_2^2}}}{\sigma_2 \sqrt{2\pi}} I = \sum_{\mu \in I} \frac{e^{-\frac{(i - \mu_1)^2}{2 \sigma_1^2}}}{\sigma_1 \sqrt{2\pi}} \mu_2 - \frac{e^{-\frac{(i - \mu_1)^2}{2 \sigma_2^2}}}{\sigma_2 \sqrt{2\pi}} \mu_2
\end{gather}
Where $i$ is a positive integer which denotes to the age of distribution, I denotes to the set of the raw age distribution with a 10-years age sensitivity, $\sigma_1, \sigma_2$ denotes to two positive constants, $\mu = \vect   {\mu_1, \mu_2}$ is a tuple with the median age of the age band and percentage distribution at 1st and 2nd entry, respectively. We then integrated the age of each six age bands.

\subsubsection{Prevalence calibration}

\hl{Source} suggested that there are significant underascertainment of the number of cases and deaths of COVID-19 in Canada. \cite{ref2} \cite{ref3} We calculated the average of the underascertainment ratio of \pendingdata and assumed the number to be constant throughout the pandemic.

\subsubsection{Vaccine immunity level estimation}

\hl{Source} suggested that the waning effect of COVID-19 vaccine efficacy.

\bibliographystyle{plain}

\begin{thebibliography}{99}
\bibitem{ref1}{Public Health Ontario. Ontario COVID-19 Data Tool. Datasets - ontario data catalogue. Retrieved August 21, 2022, from https://data.ontario.ca/en/dataset?groups=2019-novel-coronavirus}

\bibitem{ref2}{Bolotin, S., Tran, V., Deeks, S. L., Peci, A., Brown, K. A., Buchan, S. A., Ogbulafor, K., Ramoutar, T., Nguyen, M., Thakkar, R., DelaCruz, R., Mustfa, R., Maregmen, J., Woods, O., Krasna, T., Cronin, K., Osman, S., Joh, E., $\&$amp; Allen, V. G. (2021). Assessment of population infection with SARS-COV-2 in Ontario, Canada, March to June 2020. Eurosurveillance, 26(50). https://doi.org/10.2807/1560-7917.es.2021.26.50.2001559}

\bibitem{ref3}{Anand, S. S., Arnold, C., Bangdiwala, S. I., Bolotin, S., Bowdish, D., Chanchlani, R., de Souza, R. J., Desai, D., Kandasamy, S., Khan, F., Khan, Z., Langlois, M.-A., Limbachia, J., Lear, S. A., Loeb, M., Loh, L., Manoharan, B., Nakka, K., Pelchat, M., … Wahi, G. (2022). Seropositivity and risk factors for SARS-COV-2 infection in a South Asian community in Ontario: A cross-sectional analysis of a prospective cohort study. CMAJ Open, 10(3). https://doi.org/10.9778/cmajo.20220031 }
\end{thebibliography}

\end{document}